---
title: "Analytical Plan for Incidence rates of scorpion accidents in Urban Planning Areas of Americana/SP in 1998-2018"
subtitle: 'DOCUMENT: SAP-2021-007-JB-v01'
author: '**From:** Felipe Figueiredo **To:** José Brites'
date: 2021-10-21
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    reference_docx: misc/style_SAP_en.docx
    toc: yes
# toc-title: "Sumário"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(scipen = 999)
library(pander)
library(knitr)
panderOptions('table.style', 'rmarkdown')
panderOptions('table.split.table', 160)
```

---

\newpage

**Analytical Plan for Incidence rates of scorpion accidents in Urban Planning Areas of Americana/SP in 1998-2018**

**Document version**

```{r history}
Version <- c("01")
Changes <- c("Initial version")
history <- cbind(Version, Changes)
colnames(history) <- c("Version", "Alterations")
# pander(history, split.cells = 70)
kable(history)
```

---

# Abbreviations

<!-- - d: Cohen's D effect size -->
- IQR: interquartile range
- CI: confidence interval
- IRR: Incidence Rate Ratio
- SD: standard deviation
- UPA: Urban planning area

# Context

## Objectives

Describe the incidence rates of scorpion accidents in ten Urban Planning Areas of Americana/SP between 1998 and 2018.

## Hypotheses

N/A

## Study design

Ecological study, with unit of analysis as the Urban Planning Area (municipal geographic regions).

# Data

```{r cleaning, include=FALSE}
source("scripts/input.R", encoding = 'UTF-8')
```

## Raw data

Dataset with number of accidents and population for each UPA and year between 1998 -- 2018.

## Analytical dataset

In order to enhance the numerical performance of the statistical models the year was recentered with a time index starting at 1, and covering all years in the study period.

All variables in the analytical set were labeled according to the raw data provided and values were labeled according to the data dictionary for the preparation of production-quality results tables and figures.

After the cleaning process `r ncol(analytical)` variables were included in the analysis with `r nrow(analytical)` observations.
Table 1 shows the structure of the analytical dataset.

```{r analytical data}
analytical_mockup %>%
  pander(caption = "**Table 1** Analytical dataset structure after variable selection and cleaning.")
```

The analytical dataset will be included in the private version of the report, and will be omitted from the public version of the report.

# Study variables

## Primary and secondary outcomes

Primary outcome will be defined as the incidence rate of scorpion accidents in each APU.
For each APU the incidence rates will be computed as the ratio between the number of incident accidents over the population.
Incidence rates will be reported per 10000 persons.

**Specification of outcome measures** (Zarin, 2011):

1. (Domain)
2. (Specific measurement)
3. (Specific metric)
4. (Method of aggregation)

**Primary outcome**

## Covariates

Incidence rates will be adjusted by year and population.

# Statistical methods

## Statistical analyses

### Descriptive analyses

<!-- The epidemiological profile of each study group -->
<!-- groups -->
<!-- will be described at baseline. -->
<!-- Demographic and clinical variables will be described as -->
The characteristics of the UPAs will be described at the year 2018, which will be hypothesized to be representative of static characteristics.
All variables collected will be described as 
mean (SD),
median (IQR) and range
or as counts and proportions (%), as appropriate.
Distributions will be summarized in tables and visualized in exploratory plots.

### Inferential analyses

<!-- All comparisons between groups will be performed as univariate analyses. -->
<!-- Continuous variables will be compared between groups with the -->
<!-- independent t test with Welch correction. -->
<!-- paired t test with Welch correction. -->
<!-- Wilcoxon test. -->
<!-- Differences in distribution of categorical variables will be assessed with the -->
<!-- Fisher exact test. -->
<!-- chi-square test. -->
<!-- McNemar test. -->
All inferences will be derived from the statistical model.

### Statistical modeling

A multiple Poisson regression will be specified to explore the relations between the accident incidence rates and the UPAs between 1998 and 2018 (ten UPAs, 21 years, 210 observations).
The unit of analysis for this study is the UPA, and the incidence rate of accidents will be estimated for each UPA and for each year.
The population of the UPA was considered as a proxy for the exposure to accidents.
This was incorporated in the model as the offset to compute the incidence rates and allowed to vary over time.

To account for increases in incident accidents an interaction term will be included for year and UPA, which will allow the incidence rates to increase independently.

<!-- Since all groups of UPA x year yielded non-zero incident cases the model was adjusted using the Positive Poisson distribution to prevent loss of precision of non-zero incidence rates. -->

## Significance and Confidence Intervals

All analyses will be performed using the significance level of 5%.
All significance hypothesis tests and confidence intervals computed will be
two-tailed.
<!-- left-tailed. -->
<!-- right-tailed. -->

## Study size and Power

```{r}
# n <- analytical %>%
#   count(local) %>%
#   pull(n)
# n1 <- n[1]
# n2 <- n[2]
# total <- sum(n)
# d <- pwr::pwr.t2n.test(n1 = n1, n2 = n2, power = .8, sig.level = .05)$d %>%
#   format.float()
# d <- pwr::pwr.t.test(n = total/2, power = .8, sig.level = .05)$d %>%
#   format.float()
```

N/A

## Statistical packages

This analysis will be performed using statistical software `R` version `r getRversion()`.

# Observations and limitations

# References

- **SAR-2021-007-JB-v01** -- Incidence rates of scorpion accidents in Urban Planning Areas of Americana/SP in 1998-2018
<!-- - Cohen, J. (1988). Statistical power analysis for the behavioral sciences (2nd Ed.). New York: Routledge. -->
- Zarin DA, et al. The ClinicalTrials.gov results database — update and key issues. N Engl J Med 2011;364:852-60 (<https://doi.org/10.1056/NEJMsa1012065>).
- Gamble C, et al. Guidelines for the Content of Statistical Analysis Plans in Clinical Trials. JAMA. 2017;318(23):2337–2343 (<https://doi.org/10.1001/jama.2017.18556>).

# Appendix

This document was elaborated following recommendations on the structure for Statistical Analysis Plans (Gamble, 2017) for better transparency and clarity.

## Availability

Both this analytical plan and the corresponding analysis report (**SAR-2021-007-JB-v01**) can be downloaded in the following address:

<https://philsf-biostat.github.io/SAR-2021-007-JB/>

<!-- The client has requested that this analysis be kept confidential. -->
<!-- Both this analytical plan and the corresponding analysis report (**SAR-2021-007-JB-v01**) are therefore not published online and only the title and year of the analysis will be included in the consultant's Portfolio. -->
<!-- The portfolio is available at: -->

<!-- <https://philsf-biostat.github.io/> -->
